---
title: 'Social Networking profile words analysis: Clustering with H2o package and
  base R'
author: "Varun Katyal"
date: "08/01/2021"
site: bookdown::bookdown_site
output: bookdown::gitbook
---

# Loading required libraries and read dataset
```{r message=FALSE}
library(data.table)
library(mltools)
library(h2o)
library(dplyr)
library(ggplot2)


```

## importing csv file 
```{r}
sns <- read.csv('data.csv',stringsAsFactors =T )
str(sns)
```

# Exploratory Data Analysis

## distribution of gender
```{r}
sns %>% group_by(gender) %>% summarise(number=n())
```
* High proportion of females compared to males

## Distribution of No. of Friends
```{r}
ggplot(data = sns,aes(x=friends))+geom_density()+ xlim(c(0,200))+ggtitle("Distribution of Number of Friends")

```
* Most have friends from 1-50, while 100 or more is rare




## converting grad year into factor variable
```{r}
sns$gradyear <- factor(sns$gradyear)
```

## converting categorical variable into numeric(one-hot encoding) using mltools package 
## removing rows containing NAs
```{r message=FALSE, warning=FALSE}
sns <- data.table(sns)
sns_new <- sns[complete.cases(sns),]
sns_new <- one_hot(sns_new,sparsifyNAs = T)
str(sns_new)
```

## Removing age outliers
```{r}
sns_new <- sns_new %>% filter(age>=13 & age <20)
```

##  normalizing Profile's interest variables

```{r}
sns_int <- as.data.frame( apply(sns_new[,9:44],2,scale))
str(sns_int)
```


# Modelling

## initializing h2o session
```{r}

h2o.init()
```


## clustering based on interest
## creating h2o data frame
```{r}
cnames <- colnames(sns_int)
clus.hex = as.h2o(sns_int)
```

## K- clusters from 3 to 30
```{r}
p <- seq(3,30,3)
```


## Starting with 6 cluster
```{r}
mdl <- h2o.kmeans(training_frame = clus.hex,x=cnames,k=6)
mdl
```

## Plotting Elbow plot and finding optimum no. of K

```{r}
val <- sapply(p, function(i){
  mdl <- h2o.kmeans(training_frame = clus.hex,x=cnames,k=i)
  h2o.tot_withinss(mdl)
  
})
plot(p,val,type = "l",main="Elbow Plot",xlab = "K",ylab = "Total Within Sum of Squares")

```

* Bend around K = 18
* Evaluating Clustering with h2o package and base R


## Re-runing K-mean algorithm with 18 cluster
```{r}
mdl18 <- h2o.kmeans(training_frame = clus.hex,x=cnames,k=18)
mdl18
```

### With base R
* Cluster Sizes
* Within Sum of Squares

```{r}
km <- kmeans(sns_int,18)
size <- km$size
wss <- km$withinss
```


```{r}
size
```


```{r}
wss
```

## Goodness of Fit
* Ratio of between Sum of Squares and Total sum of squares
* Ideally  clustering that has the properties of internal cohesion and external separation, i.e. the ratio should approach 1.

```{r}
h2o.betweenss(mdl18)/h2o.totss(mdl18)
```

* Ratio is 0.24. Does not indicate good clustering
* With some very small cluster sizes indicates lots of outliers in different interest variables 
* Also huge difference between largest and smallest cluster show, it may be large group of teens share similar interests

## Cluster labelling in dataset
```{r}
predictions <- h2o.predict(mdl18, clus.hex)

sns_new$clust <- factor(as.vector(predictions))

```

## Examining proportion of females in each cluster
```{r message=FALSE}
fprop <- sns_new %>% group_by(clust) %>% summarise(props=mean(gender_F),size=n())
fprop

```



## Average no. of friends in each cluster
```{r}
avgf <- sns_new %>% group_by(clust) %>% summarise(avgf=mean(friends),size=n())
avgf
```


## Lets examine profile with high no. of drugs and death words
```{r}
hdd <- sns_new %>% filter(death>5 | drugs>5)
hdd

```
* Most of them belong to cluster 8 and 12




